Experimental verification

### 2018-05-07 patent provisional/Experimental verification/Experimental verification.tex
patent provisional -- first draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
%\numberwithin{figure}{section}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}[section]
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{claim}[thm]{\protect\claimname}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}

\newtheorem*{lem*}{Lemma}

\theoremstyle{remark}
\newtheorem{rem}[thm]{\protect\remarkname}
\theoremstyle{plain}
\newtheorem{corollary}[thm]{\protect\corollaryname}
\theoremstyle{plain}
\newtheorem{proposition}[thm]{\protect\propositionname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\usepackage{slashbox}

\usepackage{babel}
\providecommand{\claimname}{Claim}
\providecommand{\definitionname}{Definition}
\providecommand{\lemmaname}{Lemma}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\providecommand{\corollaryname}{Corollary}
\providecommand{\propositionname}{Proposition}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{{\textsf{SNR}}}

@@

\title{Toward single particle reconstruction without particle picking}

@@

\begin{abstract}
	Here comes the abstract
\end{abstract}

\section{Introduction}

Single particle reconstruction (SPR) using cryo--electron microscopy (cryo--EM) is an innovative technology for reconstructing the 3-D structure of macromolecules. In recent years, structures
of many molecules, previously regarded as insurmountable, are now being reconstructed to near-atomic resolution; see for instance~\cite{kuhlbrandt2014resolution,bartesaghi20152,smith2014beyond}. This technological advancement was recognized by the 2017 Nobel Prize in Chemistry~\cite{nobel}.

In a cryo--EM experiment, multiple biological samples of the (ideally) same molecule are rapidly frozen in a thin layer of vitreous ice. Within the ice, the molecules are randomly oriented and positioned. Then, the microscope produces a 2D tomographic projection image, called a \emph{micrograph}, of the multiple samples embedded in the ice. The micrograph is dominated by noise due to the small electron doses that
can be applied to the specimen without causing radiation damage.
The cryo--EM problem is to estimate the structure of the molecule from the micrograph (or, typically, several micrographs).

\paragraph{Particle picking.}
All contemporary methods in the field split the reconstruction procedure to several subroutines. 
The first step in the pipeline is the so-called \emph{particle picking}, in which one aims to detect the 2D tomographic projections of the samples from the noisy micrograph. The output of ideal particle picker is a series of 2D images, each containing one centered tomographic projection associated with an unknown viewing direction and an in-plane rotation. This series of images 
is then used to estimate the structure.

Many automatic and semi-automatic methods for particle picking have been proposed, based on edge detection, template matching and deep learning; see for instance~\cite{harauz1989automatic,ogura2004automatic,zhu2016deep,frank1983automatic,scheres2015semi,heimowitz2018apple}. 
However, most of these procedures are prone to \emph{model bias}. For instance, in the popular framework of RELION~\cite{scheres2015semi}, the user manually marks hundreds of spots on the micrograph, believed to contain projections. 
Therefore, the algorithm's performance depends on the prior assumptions of the users about the particle's structure; the same holds true for deep learning based approaches which require constructing labeled sets of data.
Other methods use disks or differences of Gaussians as templates~\cite{langlois2014automated,voss2009dog}.
Nowadays, it also still popular to pick particles manually. This method, while it exploits the researcher's experience, is both tedious and subject to model bias.

The significance of the model bias in cryo--EM was stressed in~\cite{shatsky2009method,henderson2013avoiding} and demonstrated by the ``Einstein from noise'' example.
In this experiment, the image of Einstein is correlated with multiple images of pure noise. The noisy images are then aligned to Einstein's image using cross-correlation and  averaged. 
In~\cite{shatsky2009method}, it was shown that using merely 1000 noise images, the averaged image is remarkably similar to the template image of 
Einstein, rather than to pure noise. 
%Therefore, such a simple algorithm is biased towards its model which is an undesirable property. 
In the context of cryo--EM, this example demonstrates how our prior assumptions about the particle may influence the reconstructed structure.

Besides the model bias, particle pickers have several limitations due to the high noise level, in particular for small particles and in low contrast. 
As a result, the projections in the 2D images are typically not centered; including the correct centers as parameters dramatically increases the complexity of reconstruction algorithms.
%In addition, the information from particles that are too close to each other is usually neglected. Hence, valuable information that can be harnessed is omitted.

\paragraph{Simplified model for SPR} In this paper, we present a preliminary study that aims to show that estimating a particle 
directly from the micrograph, without identifying the location of individual particle images in the micrograph, might be possible.

In particular, we consider the toy problem of estimating a set of signals $x_1,\ldots,x_K$ from their multiple occurrences in unknown  locations in a noisy data sequence $y$. Here, $y$ plays the role of the micrograph and the $K$ signals can be thought of as $K$ different viewing directions of the particle, each of which  appears multiple times in the micrograph. 
A precise mathematical formulation of the model and the estimation problem is provided in Section~\ref{sec:methods}.
To be clear, we do not consider here many prominent features of real SPR experiments and do not aim to reconstruct any 3-D structure; instead, we solve a simpler problem that we believe captures key elements of the true SPR problem. We also mention in passing that similar models emerge in different scientific fields, such as spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

Figure~\ref{fig:micro_example} shows a simple example of our problem with  one signal ($K=1$). 
Each panel demonstrates an image of size $250\times 250$, containing 4 occurrences of an Einstein image of size $50\times 50$. 
The left panel shows the data without noise. In the middle panel,  i.i.d.\ Gaussian noise with standard deviation $\sigma=0.5$  was added, however,  Einstein is still easily identified. In this scenario it is likely that standard detection algorithms can locate the copies of the signal within the micrograph. In the right panel, we show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$.
Clearly, identifying the locations of Einstein in this micrograph is much more challenging. In fact, it can be shown that in the low signal--to--noise (\SNR) regime, 
detection  of individual image occurrences is impossible---even if the true image is known---and therefore particle picking is impossible.  This phenomenon is explained in detail in Section~\ref{sec:related_literature}.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of data sequences (micrographs) of size $250\times 250$ with additive i.i.d./ Gaussian noise with variance $\sigma^2$. Each micrograph contains four occurrences of a $50 \times 50$ image of Einstein.}	
\end{figure}

\paragraph{Autocorrelation analysis.}
We now return to the general problem of estimating $x_1,\dots,x_K$ from the noisy observation $y$.
In order to recover the signals $x_1,\dots,x_K$ from the data, we use autocorrelation analysis.
In a nutshell, the method consists of two stages. First, we estimate a mixture (i.e., linear combination) of the low--order autocorrelation functions of the signals from the data. These quantities can be estimated, to any desired accuracy, if individual occurrences are separated by the maximum signal size and each signal appears sufficiently many times in the data. There is no need to detect individual occurrences.
In the second stage, the signals are estimated from the mixed autocorrelations using a a nonconvex least-squares (LS) algorithm or a phase retrieval algorithm; see Section~\ref{sec:methods}.
This method requires only one pass over the data and can recover the signals in any $\SNR$, if the signals appear enough times in the data. As a side note, we mention that expectation-maximization---a popular framework in SPR---is intractable for this problem (see Appendix~\ref{sec:theory}  for more details).

\section{Results}

%\TODO{Here we should say that we can afford to work only with non-biased terms}

\TODO{Revise to explain $K = 1$ experiment first, then explain $K = 3$.}
\TODO{TB: we should remove the ``long estimation" from the plots}

In this section we show numerical results, while leaving the specific details of the experiments for Section~\ref{sec:methods}. In the first experiment, we estimated Einstein's image from micrographs with $\sigma=3$ . An example for a micrograph appears in the right panel of Figure~\ref{fig:micro_example}. As initial guess, we picked Newton image. If the algorithm was prone to model bias, we would expect to get as an output an image that resembles Newton, similarly to the ``Einstein from noise'' effect. Nonetheless, the more data we collect, the better the reconstruction.\TODO{We have a movie in the supplementary material.}

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Newton}
		\caption{Newton  (model)}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_100}
		\caption{$P =10^2$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e3}
		\caption{$P =10^3$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e4}
		\caption{$P = 10^4$}
	\end{subfigure}%
	\caption{\label{fig:Einst_example} Recovery of Einstein image. The algorithm is initialized by the image of Newton. We show estimations of Einstein using $P=10^2,10^3$ and $10^4$ number of micrographs, each contains $860$ image occurrences on average. The normalized recovery error is $0.555$, $0.3988$ and $0.2907$, respectively. The normalized error in recovering the power spectrum was $69.1\times10^{-3} $, $21.7\times10^{-3} $ and $6.8\times10^{-3}$, respectively. \TODO{This image will be more impressive in the future}}	
\end{figure}

Here we should describe the 1-D experiments.

\begin{figure}[t]
	\centering
	\includegraphics[scale=0.6]{XP_1D_homogeneous_new_ROI/progressive_n10000000000_107521}
	\caption{\TODO{...} \TODO{This figure is with new ROI method based on loss functions}}
	\label{fig:1Dhomosignals}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\linewidth]{XP_1D_homogeneous_new_ROI/progressive_RMSE_n10000000000_107521}
	\caption{\TODO{...} \TODO{This figure is with new ROI method based on loss functions}}
	\label{fig:1DhomoRMSE}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\linewidth]{XP_1D_heterogeneous/data_example_heterogeneous_n_24600000000_604392_fig1}
	\caption{\TODO{Thin yellow is ground truth; blue is ROI of first optimization; red is final estimation. Relative error for blue is .37, and .13 for red. Individual relative errors of the red estimates: 0.0239393 / 0.208925 / 0.0335956}}
	\label{fig:1Dheterosignals}
\end{figure}

\section{Discussion}

All current SPR algorithms start with a particle picking stage which is prone to model bias. 
In this paper, we propose to examine the possibility to bypass particle picking and estimate the particle directly from the micrograph. To this end, we examine a simplified model that share some similarities with the SPR problem. The proposed reconstruction method is based on autocorrelation analysis, 
somewhat similar to a technique proposed by Kam which can be used for ab inito modeling~\cite{kam1980reconstruction,levin20173d,singer2018mathematics}. 
That being said, the SPR model is far more complicated thans the model presented here. In a future research, we hope to bridge this gap.

Our results rely on two core assumptions that are not necessarily met by in an SPR experiment.  
First, we modeled the background information as i.i.d.\ additive noise. In practice, the background information may be structured or depend on the signal. However, methods based on autocorrelation estimation are quite flexible and in many cases can tolerate mismatch of the noise statistical model.

In addition, we assumed that the signal occurrences are all separated, see~\eqref{eq:spacing}. 
This separation can be induced by careful experiment design \TODO{???}. 
If the signals are not separated, one can introduce a new variable that represents the distribution of the spacing between signal occurrences and then compute explicitly the relation between the autocorrelation of the data and the signals. It is yet to be studied under what conditions on this distribution one can estimate the signals.

%The first entry $p[1]$ will represent the probability that two consecutive signals are separated by only one entry, $p[2]$ the probability for spacing of two entries and so on. Using this auxiliary variable $p$, one can write explicitly the relation between the autocorrelation functions of the data and those of the signal in a similar way to~\eqref{eq:data_ac}. 
%An interesting question is  under what conditions on $p$ and the signals, one can estimate the signals from the data.

\section{Methods} \label{sec:methods}

@@

Let $x_1,\ldots,x_K\in\RL$ be the sought signals and let $y\in\RN$ be the data. 
The forward model can be posed as a mixture of \emph{blind deconvolution} problems between binary signals and the target signals $x_i$:
\begin{equation} \label{eq:model}
y = \sum_{i=1}^K x_i\ast s_i + \varepsilon,\quad \varepsilon\sim\mathcal{N}(0,\sigma^2 I).
\end{equation}
%We model the background information as i.i.d.\ Gaussian noise with zero mean and $\sigma^2$ variance. 
The nonzero values of each  $s_i\in\{0,1\}^N$ determine the position of the occurrences of the corresponding $x_i$. 
We denote the set of these nonzero values by 
 $\mathcal{S}_i$ and its cardinality by $\vert \mathcal{S}_i\vert = M_i$. 
By assuming that all $\mathcal{S}_i$'s are disjoint,  we let $s = \sum_{i=1}^Ks_i$, $\mathcal{S} = \bigcup_{i=1}^{K} \mathcal{S}_i$ and  $\vert \mathcal{S}\vert :=M =  \sum_{i=1}^{K}M_i$.  Neither the $M_i$'s nor $M$ are assumed to be known. Literature survey on blind deconvolution and related problems is given in Appendix~\ref{sec:related_literature}.

In order to estimate the mixture of autocorrelations, we assume that the support of $s$ is not clustered. In particular,
\begin{equation} \label{eq:spacing}
\text{For all }  i,j\in\mathcal{S}, \, i\neq j,  \quad  \vert i-j \vert\geq L-1.  
\end{equation}
The goal of the problem is to estimate $x_1,\ldots,x_K$ from $y$.

%
%\section{Autocorrelation analysis}   \label{sec:autocorrelation}
%
%Our method for estimating the signals is composed of two stages. 
%First, we use the autocorrelation functions of the data to estimate a mixture (i.e., linear combination) of the $K$ signal's autocorrelations. The mixed autocorrelation can be estimated to any accuracy, in any $\SNR$ level, if $M$ is large enough and the spacing condition~\eqref{eq:spacing} is met. Then, we  use a nonconvex LS  to estimate the signals from their mixed autocorrelations. 
%In this section, we elaborate on the autocorrelation functions and their estimations, while the precise recovery procedure, based on nonconvex optimization, will be discussed in detail in the next section.

\paragraph{Aperiodic autocorrelation functions}
%\subsection{Aperiodic autocorrelation functions} \label{sec:aperiodic_ac}

For the purpose of this paper, we need the first three (aperiodic) autocorrelation functions. The first-order autocorelation is the mean of the signals. For  
$z\in\RL$ and $k\geq 2$, the autocorrelation of order $k$ is defined for any integer shifts $\ell_1, \ldots, \ell_{k-1}$ by
\begin{align}
	a_z^k[\ell_1,\ldots,\ell_{k-1}]  & = \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\ldots z[i+\ell_{k-1}],
	\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the bounds $0, \ldots, L-1$ is zero-padded, as usual.
Explicitly, the first three autocorrelations are
\begin{align} 
	a_z^1 & = \sum_{i=0}^{L-1} z[i], \nonumber\\
	a_z^2[\ell] & = \sum_{i = \max\{0, -\ell\}}^{L-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
	a_z^3[\ell_1,\ell_2] & = \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{L-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
Note that the autocorrelation functions are symmetric so that $a_z^2[\ell] = a_z^2[-\ell]$ and $$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].$$
%Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.

\paragraph{Estimating the autocorrelation function of a single signal}

We first consider the problem of estimating the autocorrelations of a single signal from the data. 
The main principles carry through for $K>1$ as will be shown next.

%In order to estimate the autocorrelations of the signal, we first compute the first $L$ entries of the data's autocorrelations. 
For the purpose of the analysis, we consider  the asymptotic regime where $M,N\to\infty$, while preserving fixed ratio. 
Specifically, we define the ratio of the measurement occupied by  the signal as
\begin{equation}
\gamma = \frac{M L}{N}.
\end{equation}
Under the spacing constraint~\eqref{eq:spacing}, we have $\gamma\leq\frac{L}{2L-1}\approx 1/2$.

The main pillar of this work is the following simple observation.
If the support signal $s$ satisfies the spacing constraint~\eqref{eq:spacing}, then the first $L$ entries of the data autocorrelations converge 
to a scaled, biased, version of the signal's autocorrelation:
\begin{align}
\lim_{N\to\infty} a_y^1 & = \gamma a_{x}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \gamma a_{x}^2[\ell] +\sigma^2\delta[\ell], \label{eq:data_ac_k1} \\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2] + \sigma^2\gamma a_{x}^1 \big(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]\big), \nonumber
\end{align}
for $\ell,\ell_1,\ell_2=0,\ldots L-1$, and where $\delta$ denotes the Kronecker delta function. 
%These relations are proven in Appendix~\ref{sec:autocorrelation_computation}.

\paragraph{Estimating the autocorrelation function of a multiple signals}

As before, we consider  the asymptomatic regime where $M_1,\ldots,M_K,N\to\infty$, while preserving fixed ratios
\begin{align}
	\gamma_k = \frac{M_k L}{N}, \quad \gamma = \sum_{k=1}^K\gamma_k.
\end{align}
If the support $s$ satisfies the spacing constraint~\eqref{eq:spacing}, then one can estimate the mixture of the $K$ signals'  autocorrelations, similarly to~\eqref{eq:data_ac_k1}: 
\begin{align}
\lim_{N\to\infty} a_y^1 & = \sum_{k=1}^K\gamma_k a_{x_k}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \sum_{k=1}^K\gamma_k a_{x_k}^2[\ell] +\sigma^2\delta[\ell],  \label{eq:data_ac}\\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \sum_{k=1}^K\gamma_k a_{x_k}^3[\ell_1,\ell_2] + \sigma^2\left(\sum_{k=1}^K\gamma_k a_{x_k}^1\right)(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]), \nonumber
\end{align}
This relation is proven in  Appendix~\ref{sec:autocorrelation_computation}.

\paragraph{Numerical experiments details}

\TODO{We may want to split it into two sections. Here we could give a short description with the LS estimator and the RRR. In the supplementary material we can provide all other details. }

For the 1D experiment, we fix $K = 3$ signals of length $L = 21$, as depicted in Figure~\TODO{ref}. Following the data model described in Section~\TODO{ref}, we generate an observation $y$ of length $24.6 \cdot 10^9$. Each of the three signals appears, respectively (and approximately) $300 \cdot 10^6$, $200 \cdot 10^6$ and $100 \cdot 10^6$ times in $y$, such that at least $L-1$ zeros separate two occurrences of any signals. This is done by randomly selecting $600 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint and locations picked so far. For each placement, one of the three signals is picked at random proportionally to the desired number of occurrences of each. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The SNR of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/12.
%Given the scaling of the target signals (whose median absolute value for the entries is at or below one),
This is enough noise to make cross-correlations of $y$ even with the true signals display peaks at random locations, uninformative of the actual locations of the signal occurrences. Thus, we contend that it would be difficult for any algorithm to locate the signal occurrences, let alone to classify them according to which signal appears where.

\TODO{TB: I would place the equation counting argument in the theory section in the paragraph of open questions (I marked th place)}
Given the observation $y$, we proceed to compute the moments. The first-order moment is straightforward. For second-order moments, notice from equation~\eqref{eq:data_ac} that $a_y^2[\ell]$ suffers no bias for $\ell$ in $1$ to $L-1$. Thus, we omit $\ell = 0$, which has the practical effect that we need not know $\sigma$ to estimate the moments. Likewise, for third-order moments, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant moments for our purpose \TODO{Do we want to include a figure to explain why that is?}, and we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid biased elements---there are $\frac{(L-1)(L-2)}{2}$ remaining moments. As a result, it is unnecessary to estimate $\sigma$. We have \TODO{TB: We may want to put this calculation is the last paragraph of Section 3.1 }
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
moments in total. In practice, these are computed on disjoint segments of $y$ of length $100\cdot10^6$ and added up, without correction for the junction points. Segments are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
%\numberwithin{equation}{section}
%\numberwithin{figure}{section}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}[section]
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{claim}[thm]{\protect\claimname}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}
%
%\newtheorem*{lem*}{Lemma}
%
%\theoremstyle{remark}
%\newtheorem{rem}[thm]{\protect\remarkname}
%\theoremstyle{plain}
%\newtheorem{corollary}[thm]{\protect\corollaryname}
%\theoremstyle{plain}
%\newtheorem{proposition}[thm]{\protect\propositionname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\usepackage{slashbox}

\usepackage{babel}
%\providecommand{\claimname}{Claim}
%\providecommand{\definitionname}{Definition}
%\providecommand{\lemmaname}{Lemma}
%\providecommand{\remarkname}{Remark}
%\providecommand{\theoremname}{Theorem}
%\providecommand{\corollaryname}{Corollary}
%\providecommand{\propositionname}{Proposition}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
%\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
%\newcommand{\hx}{\hat{x}} 
%\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{{\textsf{SNR}}}

@@

\title{Experimental verification}

@@

In this note, we present a preliminary study that aims to show that estimating a particle 
directly from the micrograph, without identifying the location of individual particle images in the micrograph, might be possible. 
In particular, we consider the problem of estimating a set of signals $x_1,\ldots,x_K$ from their multiple occurrences in unknown  locations in a noisy data sequence $y$. Here, $y$ plays the role of the micrograph and the $K$ signals can be thought of as $K$ different viewing directions of the particle, each of which  appears multiple times in the micrograph. 
To be clear, we do not consider here many prominent features of real SPR experiments and do not aim to reconstruct any 3-D structure; instead, we solve a simpler problem that we believe captures key elements of the true SPR problem.

Figure~\ref{fig:micro_example} shows a simple example of our problem with  one signal ($K=1$). 
Each panel demonstrates an image of size $250\times 250$, containing 4 occurrences of an Einstein image of size $50\times 50$. 
The left panel shows the data without noise. In the middle panel,  i.i.d.\ Gaussian noise with standard deviation $\sigma=0.5$  was added, however,  Einstein is still easily identified. In this scenario it is likely that standard detection algorithms can locate the copies of the signal within the micrograph. In the right panel, we show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$.
Clearly, identifying the locations of Einstein in this micrograph is much more challenging. In fact, it can be shown that in the low signal--to--noise (\SNR) regime, 
detection  of individual image occurrences is impossible---even if the true image is known---and therefore particle picking is impossible~\cite{aguerrebere2016fundamental}.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_clean}
		\caption{$\sigma = 0$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s05}
		\caption{$\sigma = 0.5$}
	\end{subfigure}
	\begin{subfigure}[h]{0.33\textwidth}
		\centering
		\includegraphics[scale=0.5]{micrograph_Einstein_example_s3}
		\caption{$\sigma = 3$}
	\end{subfigure}
	\caption{\label{fig:micro_example} Example of data sequences (micrographs) of size $250\times 250$ with additive i.i.d.\ Gaussian noise with variance $\sigma^2$. Each micrograph contains four occurrences of a $50 \times 50$ image of Einstein.}	
\end{figure}

@@

To be concrete, we formulate the problem for one-dimensional signals. The extension to multi-dimensional signals is straight-forward.

Let $x_1,\ldots,x_K\in\RL$ be the sought signals and let $y\in\RN$ be the data. 
The forward model can be posed as a mixture of {blind deconvolution} problems between binary signals and the target signals $x_i$:
\begin{equation} \label{eq:model}
y = \sum_{i=1}^K x_i\ast s_i + \varepsilon,\quad \varepsilon\sim\mathcal{N}(0,\sigma^2 I).
\end{equation}
The nonzero values of each  $s_i\in\{0,1\}^N$ determine the position of the occurrences of the corresponding $x_i$; see Figure~\ref{fig:micro_example} for an example. We assume that the nonzero values of $s$ are separated by at least $L-1$ entries. In the multi-dimensional case, this separation is imposed in each axis.

\paragraph{Autocorrelation analysis.}
In order to recover the signals $x_1,\dots,x_K$ from the data, we use autocorrelation analysis.
In a nutshell, the method consists of two stages. First, we estimate a mixture (i.e., linear combination) of the low--order autocorrelation functions of the signals from the data. Then, we estimate the signal itself using non-convex least-squares (LS) or a phase retrieval algorithm.

\paragraph{Autocorrelation functions.}
We recall that the first-order autocorelation is the mean of the signals. For  
$z\in\RL$ and $k\geq 2$, the autocorrelation of order $k$ is defined for any integer shifts $\ell_1, \ldots, \ell_{k-1}$ by
\begin{align}
a_z^k[\ell_1,\ldots,\ell_{k-1}]  & = \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\ldots z[i+\ell_{k-1}],
\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the bounds $0, \ldots, L-1$ is zero-padded, as usual.
Explicitly, the first three autocorrelations are
\begin{align} 
a_z^1 & = \sum_{i=0}^{L-1} z[i], \nonumber\\
a_z^2[\ell] & = \sum_{i = \max\{0, -\ell\}}^{L-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
a_z^3[\ell_1,\ell_2] & = \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{L-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}

\paragraph{Esimtating the mixture of the signals' autocorrelation from the data.}
A mixture of the $K$ signals' autocorrelation functions can be estimated, to any desired accuracy, if individual occurrences are separated by  $L-1$ entries in each direction and each signal appears sufficiently many times in the data. There is no need to detect individual occurrences.

Let $M_i$ denote the number of occurrences of the $i$th signal. 
Let us denote
\begin{align}
\gamma_k = \frac{M_k L}{N}.
\end{align}
Then, under the spacing constraint discussed above, one can show that
\begin{align} 
\lim_{N\to\infty} a_y^1 & = \sum_{k=1}^K\gamma_k a_{x_k}^1, \nonumber\\
\lim_{N\to\infty} a_y^2[\ell] & = \sum_{k=1}^K\gamma_k a_{x_k}^2[\ell] +\sigma^2\delta[\ell],  \label{eq:data_ac}\\
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] & = \sum_{k=1}^K\gamma_k a_{x_k}^3[\ell_1,\ell_2] + \sigma^2\left(\sum_{k=1}^K\gamma_k a_{x_k}^1\right)(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]), \nonumber
\end{align}

\paragraph{Numerical experiments.}
In the first experiment, we estimated Einstein's image from micrographs with $\sigma=3$ . An example for a micrograph appears in the right panel of Figure~\ref{fig:micro_example}.
We used a micrograph of size $4000\times 4000$ and insert Einstein's images in random locations, while preserving the separation constraint. On average, each micrograph includes $860$ images. We then computed the first
We computed the first and second moment of the data and estimated the associated moments of the image using~\eqref{eq:data_ac}. We recall that the second moment is equivalent to the Fourier magnitude of the signal and therefore one can use standard phase retrieval algorithms. Here, we use an algorithm called relaxed-reflect-reflect (RRR)~\cite{elser2017matrix} that iterates by 
\begin{equation}
x_{\ell+1} = x_\ell  + P_2(2P_1(x_\ell) - x_\ell) - P_1(x_\ell),
\end{equation}
where $P_2(z)$ combines the Fourier phases of $z$ with the Fourier magnitudes estimated from the data, and $P_1$ zeros out the signal out of its support. 
If the Fourier magnitude are estimated exactly, then the algorithm produces the signal or its reflection through the origin. 
In  we show an example how the reconstruction error decreases with as we collect more micrographs.

 \TODO{We have a movie in the supplementary material.} 
 \TODO{To add a figure of the error as a function of the number of micrographs.}

\begin{figure}[h!]
	\centering
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_100}
		\caption{$P =10^2$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e3}
		\caption{$P =10^3$}
	\end{subfigure}%
	\begin{subfigure}[h]{0.25\textwidth}
		\centering
		\includegraphics[scale=0.4]{Einstein_1e4}
		\caption{$P = 10^4$}
	\end{subfigure}%
	\caption{\label{fig:Einst_example} Recovery of Einstein image. The algorithm is initialized by the image of Newton. We show estimations of Einstein using $P=10^2,10^3$ and $10^4$ number of micrographs, each contains $860$ image occurrences on average. The normalized recovery error is $0.555$, $0.3988$ and $0.2907$, respectively. The normalized error in recovering the power spectrum was $69.1\times10^{-3} $, $21.7\times10^{-3} $ and $6.8\times10^{-3}$, respectively.}	
\end{figure}
